---
title: "What makes loans default faster?"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyverse)
accounts <- read.csv("/Users/zhiyulin/anly503/fall2020/anly503-fall2020-a5-zlds123/data/accounts_analytical.csv")
transactions <- read.csv("/Users/zhiyulin/anly503/fall2020/anly503-fall2020-a5-zlds123/data/transactions.csv")
temp <- accounts[,c(names(accounts)[grepl( "loan" , names(accounts))], "acct_creation_date")] 
temp <- na.omit(temp)

temp$time_diff <- mapply(function(y,x) difftime(strptime(x, '%Y-%m-%d'), strptime(y, '%Y-%m-%d'), units='days'), temp$acct_creation_date, temp$loan_date)
temp$loan_date_date <- as.vector(apply(as.matrix(temp$loan_date), 1, function(x) strptime(x, "%Y-%m-%d")))
temp$account_date_date <- as.vector(apply(as.matrix(temp$acct_creation_date), 1, function(x) strptime(x, "%Y-%m-%d")))
temp$loan_default <- ifelse(temp$loan_default == TRUE, 1, 0)
temp$loan_status <- ifelse(temp$loan_status == "expired", 0, 1)
```


#### Loan parameter correlations
```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
library(ggplot2)
library(ggcorrplot)

num_temp <- temp %>% select_if(is.numeric)
r <- cor(num_temp, use="complete.obs")
ggcorrplot(r)
```

This visualization shows the correlation between every pair of loan related parameter. We observe that loan payment have a high correlation with debit_bank_transfer_loan_payment_avg_txn. 

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
b <- c(-Inf, as.numeric(summary(temp$loan_amount)[2]), as.numeric(summary(temp$loan_amount)[3]), as.numeric(summary(temp$loan_amount)[5]), Inf)
names <- c("q1", "q2", "q3", "q4")
temp$loan_amount_cat <- cut(temp$loan_amount, breaks = b, labels = names)
```

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
library(survival)
library(survminer)
sfit <- survfit(Surv(time_diff, loan_default) ~ loan_amount_cat , data=temp)
ggsurvplot(sfit, 
           conf.int=TRUE, 
           pval=TRUE,
           legend.labs=c("q1", "q2", "q3", "q4"), 
           legend.title="loan status",  
           palette=c("cornflowerblue", "indianred3", "orange", "darkgreen"), 
           title="Kaplan-Meier Curve for loan default survival",
           xlab = "Time (days)")
```

We see that loans with larger amounts are more likely to default over time.

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
sfit <- survfit(Surv(time_diff, loan_default) ~ loan_status , data=temp)
ggsurvplot(sfit, 
           conf.int=TRUE, 
           pval=TRUE,
           legend.labs=c("expired", "current"), 
           legend.title="loan status",  
           palette=c("cornflowerblue", "indianred3"), 
           title="Kaplan-Meier Curve for loan default survival",
           xlab = "Time (days)")
```

Current loans not paid off until after 650 days are significantly more likely to default than others.

```{r, echo=FALSE, warning=FALSE,message=FALSE,error=FALSE, results='hide',fig.keep='last'}
sfit <- survfit(Surv(time_diff, loan_default) ~ as.factor(loan_term) , data=temp)
ggsurvplot(sfit, 
           conf.int=FALSE, 
           pval=TRUE,
           legend.labs=c("24", "12", "60", "36", "48"), 
           legend.title="loan status",  
           palette=c("cornflowerblue", "indianred3", "orange", "darkgreen", "navy"), 
           title="Kaplan-Meier Curve for loan default survival",
           xlab = "Time (days)")
```

Loan with more than 60 months payment terms are very likely to default after 650 days compared to the others.